Static Analysis by Path Focusing

### 2011-05-04 report/report.tex
update of the report

This computation is a fixpoint computation: one updates the different sets until
we find no more elements. These sets are strictly increasing (we always add new
elements in the sets), and if we find no new elements to add, that means we have
reached a limit.

@@

The fixpoint computation algorithm consist in replacing iteratively each $X_p$ by its value
on the right hand side, until the convergence.

@@

@@

When computing the fixpoint iterations, we only consider a subset of the graph
for a while, until the convergence. In our example:
\begin{itemize}
\item Step 1:
	we only consider the path of
	the loop that is possible at the first iteration. The second path is
	ignored, and will be treated later. we compute the iterations, and we obtain
	an first invariant for this part of the graph. Then, we can do some
	narrowing steps in order to recover some precision, before adding new paths
	in the graph.
\item Step 2:
	Now, we add the new paths into our graph if they have become possible. In
	our case, the second path of the loop is now possible, so we add the path in
	the graph and compute new iterations until convergence\dots
\item Step 3: finally, the path from $p_4$ to $p_7$, so we add it into the graph and
compute the invariant for $p_7$.
\end{itemize}

@@

      if (y < 0) break;
      x++;
   }
}
\end{C}
\end{minipage}
\begin{minipage}[c]{.45\linewidth}
\includegraphics[width=8.5cm]{"images/cfg_gopan"}
\end{minipage}
\caption{Our running example and its internal representation in SSA form}
\end{figure}
\FloatBarrier

@@

@@

	\begin{definition}[Liveness\dots]
	A variable $v$ is \emph{live\dots} at the control point $p$ iff:
	\begin{itemize}
	\item 
	its value is available at $p$, meaning that there is a path from the
	point defining the variable to $p$,
	\item one of these conditions holds:
		\begin{itemize}
		\item $v$ is live in $p$.
		\item There is a variable $v'$, defined as a linear combination of other
		variables $v_1, v_2, \dots, v_n$, such that $\exists i \in \{1,\dots,n\}, v = v_i$,
		and $v'$ is live\dots in $p$.
		\end{itemize}
	\end{itemize}
	\end{definition}

	Finally, a variable is a dimension of $X_p$ iff it is live\dots and it is
	not defined as a linear combination of program variables.

>>>>

This computation is a fixpoint computation: one updates the different sets until
we find no more elements. These sets are strictly increasing (we always add new
elements in the sets), and if we find no new elements to add, that means we have
reached a limit. Abstract Interpretation gives a general framework to do such
kind of fixpoint computation.

@@

The fixpoint computation algorithm consists in replacing iteratively each $X_p$ by its value
on the right hand side, until the convergence.

@@

\section{Example}

We apply the linear relation analysis technique over the program in figure
\ref{runningexample}. The set of state where we apply widening is $\{p_1\}$: it
is sufficient, since removing $p_1$ cuts all cycles in the graph.

To each point of the control flow graph $\{p_0, p_1,\dots, p_7\}$, we attach a
convex polyhedron, whose dimensions are the numerical variables $x,y$.
For instance, at point $p_1$, the effective set of possible assignments for
$(x,y)$ is the union of the two segments $(0,0) - (51,51)$ and $(51,51) -
(102,0)$, depicted in figure \ref{result}.a. Since this union of segments is not
a convex polyhedra, the best result we can wait for after abstraction is the
polyhedra depicted in figure \ref{result}.b, which is the smallest convex
polyhedra containing the two segments.

\begin{figure}[!h]
\begin{minipage}[c]{.39\linewidth}
\centering
\begin{tikzpicture}[y=.2cm, x=.2cm,font=\footnotesize]

	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	\draw[line] (0,0) -- (8,8) -- cycle;
	\draw[line] (16,0) -- (8,8) -- cycle;

 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);

	%ticks and labels      
     		\draw [dotted](8,8) -- (-3pt,8) 
     			node[anchor=east] {$51$}; 
     		\draw [dotted] (8,8) -- (8,-3pt) 
     			node[anchor=north] {$51$}; 
     		\draw [dotted] (16,1pt) -- (16,-3pt) 
     			node[anchor=north] {$102$};

	\node[right=1.9cm] at (x axis mid) {$x$};
	\node[above=1.2cm] at (y axis mid) {$y$};

	\node[right=3cm] at (y axis mid) {$p_1$};
\end{tikzpicture}

a. During execution, $(x,y)$ can only be in one of these two segments.
\end{minipage} 
\begin{minipage}[c]{.59\linewidth}
\centering
\begin{tikzpicture}[y=.2cm, x=.2cm,font=\footnotesize]

	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	\draw[line] (0,0) -- (8,8) -- cycle;
	\draw[line] (16,0) -- (8,8) -- cycle;

	\fill[polyhedra] (0,0) -- (8,8) -- (16,0) -- cycle;

 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);

	%ticks and labels      
     		\draw [dotted](8,8) -- (-3pt,8) 
     			node[anchor=east] {$51$}; 
     		\draw [dotted] (8,8) -- (8,-3pt) 
     			node[anchor=north] {$51$}; 
     		\draw [dotted] (16,1pt) -- (16,-3pt) 
     			node[anchor=north] {$102$};

	\node[right=1.9cm] at (x axis mid) {$x$};
	\node[above=1.2cm] at (y axis mid) {$y$};

	\node[right=3cm] at (y axis mid) {$p_1$};
\end{tikzpicture}

b. The best result we can have is the polyhedra in hashed lines.
\end{minipage}
\caption{Best results we can get from the analysis}
\label{result}
\end{figure}

We apply abstract interpretation to this program. Let $X_i$ be the polyhedra
attached to point $p_i$. Initially, $X_i = \perp$ for each $i \in \{1,7\}$, and
$X_0 = \top$. The tabular in figure \ref{iterations} shows the evolution of the
polyhedron during the iterations.

After 3 iterations, we have reached an invariant. We can then apply a narrowing
iteration, that recovers some precision. Finally, at point $p_1$, we see that
the obtained result is not as precise as we could wish (Figure \ref{result}.b).
This is due to the widening operator applied at point $p_1$, and the convex hull
applied at point $p_4$.

\begin{figure}[!h]
\centering
\caption{Polyhedron attached to the different points during iterations.}
\label{iterations}
\begin{tabular}{|c||c|c|c|c|} \hline
Iteration & $X_0 \sqcup X_6$ & $X_1$ (widening) & $X_4$ & $X_6$ \\ \hline
1& 
\begin{tikzpicture}[y=.1cm, x=.1cm,font=\footnotesize]
	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	\fill[line] (0,0) circle (0.8);
	%\draw[line] (0,0) -- (8,8) -- cycle;
	%\draw[line] (16,0) -- (8,8) -- cycle;
	%\fill[polyhedra] (0,0) -- (8,8) -- (16,0) -- cycle;
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);
	%ticks and labels      
	\node[right=0.9cm] at (x axis mid) {$x$};
	\node[above=0.6cm] at (y axis mid) {$y$};
\end{tikzpicture} 
&
\begin{tikzpicture}[y=.1cm, x=.1cm,font=\footnotesize]
	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	\fill[line] (0,0) circle (0.8);
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);
	%ticks and labels      
	\node[right=0.9cm] at (x axis mid) {$x$};
	\node[above=0.6cm] at (y axis mid) {$y$};
\end{tikzpicture} 
& 
\begin{tikzpicture}[y=.1cm, x=.1cm,font=\footnotesize]
	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	\fill[line] (0,2) circle (0.8);
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);
	%ticks and labels      
     		\draw [dotted](1pt,2) -- (-3pt,2) 
     			node[anchor=east] {$1$}; 
	\node[right=0.9cm] at (x axis mid) {$x$};
	\node[above=0.6cm] at (y axis mid) {$y$};
\end{tikzpicture} 
& 
\begin{tikzpicture}[y=.1cm, x=.1cm,font=\footnotesize]
	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	\fill[line] (2,2) circle (0.8);
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);
	%ticks and labels      
     		\draw [dotted](2,2) -- (-3pt,2) 
     			node[anchor=east] { {\tiny $1$}}; 
     		\draw [dotted](2,2) -- (2,-3pt) 
     			node[anchor=north] { {\tiny $1$}}; 
	\node[right=0.9cm] at (x axis mid) {$x$};
	\node[above=0.6cm] at (y axis mid) {$y$};
\end{tikzpicture} 
\\ \hline
2&
\begin{tikzpicture}[y=.1cm, x=.1cm,font=\footnotesize]
	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	\draw[line] (0,0) -- (2,2) -- cycle;
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);
	%ticks and labels      
     		\draw [dotted](2,2) -- (-3pt,2) 
     			node[anchor=east] { {\tiny $1$}}; 
     		\draw [dotted](2,2) -- (2,-3pt) 
     			node[anchor=north] { {\tiny $1$}}; 
	\node[right=0.9cm] at (x axis mid) {$x$};
	\node[above=0.6cm] at (y axis mid) {$y$};
\end{tikzpicture} 
& 
\begin{tikzpicture}[y=.1cm, x=.1cm,font=\footnotesize]
	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	\draw[line] (0,0) -- (12,12) -- cycle;
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);
	%ticks and labels      
	\node[right=0.9cm] at (x axis mid) {$x$};
	\node[above=0.6cm] at (y axis mid) {$y$};
\end{tikzpicture} 
&
\begin{tikzpicture}[y=.1cm, x=.1cm,font=\footnotesize]
	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	\draw[line] (0,2) -- (6,8) -- cycle;
	\draw[line] (8,6) -- (14,12) -- cycle;
	\fill[polyhedra] (0,2) -- (10,12) -- (14,12) -- (8,6) -- cycle;
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);
	%ticks and labels      
	     	\draw [dotted](6,8) -- (-3pt,8) 
     			node[anchor=east,yshift=2pt] { {\tiny $51$}}; 
     		\draw [dotted] (8,6) -- (-3pt,6) 
     			node[anchor=east, yshift=-2pt] { {\tiny $50$}}; 
     		\draw [dotted](8,6) -- (8,-3pt) 
     			node[anchor=north,xshift=2pt] { {\tiny $51$}}; 
     		\draw [dotted] (6,8) -- (6,-3pt) 
     			node[anchor=north,xshift=-2pt] { {\tiny $50$}};
     		\draw [dotted](1pt,2) -- (-3pt,2) 
     			node[anchor=east] { {\tiny $1$}}; 
	\node[right=0.9cm] at (x axis mid) {$x$};
	\node[above=0.6cm] at (y axis mid) {$y$};
\end{tikzpicture} 
&
\begin{tikzpicture}[y=.1cm, x=.1cm,font=\footnotesize]
	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	%\draw[line] (2,2) -- (6,8) -- cycle;
	%\draw[line] (8,6) -- (14,12) -- cycle;
	\fill[polyhedra] (2,2) -- (12,12) -- (16,12) -- (10,6) -- cycle;
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);
	%ticks and labels      
     		\draw [dotted] (10,6) -- (-3pt,6) 
     			node[anchor=east, yshift=-2pt] { {\tiny $50$}}; 
     		\draw [dotted](10,6) -- (10,-3pt) 
     			node[anchor=north,xshift=2pt] { {\tiny $52$}}; 
     		\draw [dotted](2,2) -- (-3pt,2) 
     			node[anchor=east] { {\tiny $1$}}; 
     		\draw [dotted](2,2) -- (2,-3pt) 
     			node[anchor=north] { {\tiny $1$}}; 
	\node[right=0.9cm] at (x axis mid) {$x$};
	\node[above=0.6cm] at (y axis mid) {$y$};
\end{tikzpicture} 
\\ \hline
3& 
\begin{tikzpicture}[y=.1cm, x=.1cm,font=\footnotesize]
	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	%\draw[line] (2,2) -- (6,8) -- cycle;
	%\draw[line] (8,6) -- (14,12) -- cycle;
	\fill[polyhedra] (0,0) -- (12,12) -- (16,12) -- (10,6) -- cycle;
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);
	%ticks and labels      
     		\draw [dotted] (10,6) -- (-3pt,6) 
     			node[anchor=east, yshift=-2pt] { {\tiny $50$}}; 
     		\draw [dotted](10,6) -- (10,-3pt) 
     			node[anchor=north,xshift=2pt] { {\tiny $52$}}; 
	\node[right=0.9cm] at (x axis mid) {$x$};
	\node[above=0.6cm] at (y axis mid) {$y$};
\end{tikzpicture} 
& 
\begin{tikzpicture}[y=.1cm, x=.1cm,font=\footnotesize]
	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	\draw[line] (-4,-4) -- (12,12) -- cycle;
	\fill[polyhedra] (-4,-4) -- (12,12) -- (20,12) -- (20,-4) -- cycle;
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);
	%ticks and labels      
	\node[right=0.9cm] at (x axis mid) {$x$};
	\node[above=0.6cm] at (y axis mid) {$y$};
\end{tikzpicture} 
& 
\begin{tikzpicture}[y=.1cm, x=.1cm,font=\footnotesize]
	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	\draw[line] (-4,-2) -- (10,12) -- cycle;
	\fill[polyhedra] (-4,-4) -- (-4,-2) -- (10,12) -- (20,12) -- (20,-4) -- cycle;
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);
	%ticks and labels      
     		\draw [dotted](1pt,2) -- (-3pt,2) 
     			node[anchor=east] { {\tiny $1$}}; 
	\node[right=0.9cm] at (x axis mid) {$x$};
	\node[above=0.6cm] at (y axis mid) {$y$};
\end{tikzpicture} 
&
\begin{tikzpicture}[y=.1cm, x=.1cm,font=\footnotesize]
	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	\draw[line] (0,0) -- (12,12) -- cycle;
	\fill[polyhedra] (0,0) -- (12,12) -- (20,12) -- (20,0) -- cycle;
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);
	%ticks and labels      
	\node[right=0.9cm] at (x axis mid) {$x$};
	\node[above=0.6cm] at (y axis mid) {$y$};
\end{tikzpicture} 
\\ \hline
narrowing& 
\begin{tikzpicture}[y=.1cm, x=.1cm,font=\footnotesize]
	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	\draw[line] (0,0) -- (12,12) -- cycle;
	\fill[polyhedra] (0,0) -- (12,12) -- (20,12) -- (20,0) -- cycle;
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);
	%ticks and labels      
	\node[right=0.9cm] at (x axis mid) {$x$};
	\node[above=0.6cm] at (y axis mid) {$y$};
\end{tikzpicture} 
& 
\begin{tikzpicture}[y=.1cm, x=.1cm,font=\footnotesize]
	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	\draw[line] (0,0) -- (12,12) -- cycle;
	\fill[polyhedra] (0,0) -- (12,12) -- (20,12) -- (20,0) -- cycle;
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);
	%ticks and labels      
	\node[right=0.9cm] at (x axis mid) {$x$};
	\node[above=0.6cm] at (y axis mid) {$y$};
\end{tikzpicture} 
& 
\begin{tikzpicture}[y=.1cm, x=.1cm,font=\footnotesize]
	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	\draw[line] (0,2) -- (10,12) -- cycle;
	\draw[line] (0,2) -- (10,-2) -- cycle;
	\draw[line] (20,-2) -- (10,-2) -- cycle;
	\fill[polyhedra] (0,2) -- (10,12) -- (20,12) -- (20,-2) --
	(10,-2) -- cycle;
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);
	%ticks and labels      
     		\draw [dotted](1pt,2) -- (-3pt,2) 
     			node[anchor=east] { {\tiny $1$}}; 
     		\draw [dotted](10,-2) -- (-3pt,-2) 
     			node[anchor=east] { {\tiny $-1$}}; 
     		\draw [dotted](10,-2) -- (10,1pt) 
     			node[anchor=south] { {\tiny $51$}}; 
	\node[right=0.9cm] at (x axis mid) {$x$};
	\node[above=0.6cm] at (y axis mid) {$y$};
\end{tikzpicture} 
& 
\begin{tikzpicture}[y=.1cm, x=.1cm,font=\footnotesize]
	\node (t1) at (-5,-4) {};
	\node (t2) at (13,11) {};
	\draw[line] (2,2) -- (12,12) -- cycle;
	\draw[line] (2,2) -- (7,0) -- cycle;
	\fill[polyhedra] (2,2) -- (12,12) -- (20,12) -- (20,0) --
	(7,0) -- cycle;
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,10);
	%ticks and labels      
     		\draw [dotted](2,2) -- (-3pt,2) 
     			node[anchor=east] { {\tiny $1$}}; 
     		\draw [dotted](2,2) -- (2,-3pt) 
     			node[anchor=north] { {\tiny $1$}}; 
	\node[right=0.9cm] at (x axis mid) {$x$};
	\node[above=0.6cm] at (y axis mid) {$y$};
\end{tikzpicture} 
\\ \hline
\end{tabular}
\end{figure}

@@

When computing the fixpoint iterations, we only consider a subset of the graph
for a while, until the convergence. In our example:
\begin{itemize}
\item Step 1:
	we only consider the path of
	the loop that is possible at the first iteration. The second path is
	ignored, and will be treated later. we compute the iterations, and we obtain
	an first invariant for this part of the graph: at point $p_1$, $x =
	y \wedge x \geq 0$.
	Then, we can do some
	narrowing steps in order to recover some precision, before adding new paths
	in the graph. At point $p_1$ after narrowing, $x=y \wedge 0 \leq x \leq 51$.
\item Step 2:
	Now, we add the new paths into our graph if they have become possible. In
	our case, the second path of the loop is now possible, because $x$ can be
	equal to $51$, so we add the path in
	the graph and compute new iterations until convergence\dots. After
	convergence, at point $p_1$, $x + y - 102 \leq 0 \wedge x
	\leq y$. We then apply a narrowing step and at $p_1$, 
$x + y - 102 \leq 0 \wedge x \leq y \wedge y \geq 0$.
\item Step 3: finally, the path from $p_4$ to $p_7$ becomes possible, so we add
it into the graph and compute the invariant for $p_7$.
\end{itemize}

Finally, at $p_1$, the result we obtain is the best we can have using convex
polyhedra (see Figure \ref{result}.b). This method allows to find the
constraint $x + y - 102 \leq 0$, which is not the case in the classical abstract
interpretation technique.

@@

      if (y < 0) break;
      x++;
   }
}
\end{C}
\end{minipage}
\begin{minipage}[c]{.45\linewidth}
\includegraphics[width=8.5cm]{"images/cfg_gopan"}
\end{minipage}
\caption{Our running example and its internal representation in SSA form}
\label{IR}
\end{figure}
\FloatBarrier

As we see if Figure \ref{IR}, LLVM IR is slightly different from
our definition of control flow graph: it uses the notion of \emph{BasicBlock}.
A \emph{BasicBlock} is a series of instructions that are necessarily executed
one after the others, meaning that if the first intruction of a BasicBlock is
executed, then the rest of the instructions are necessarily executed exactly
once, in order.

In our case, a control point will be a BasicBlock of the program.

@@

	\subsection{Attachment to LLVM Internal Representation}

	We attach only one abstract value to each BasicBlock of the IR. Indeed,
	there would be a huge memory consumption if we would attach a polyhedra
	after each instruction. A BasicBlock begins with all its $\Phi$-variables,
	and ends with the other instructions. We choose to attach our abstract value
	just after the $\Phi$-variables assignments. In this way, the
	$\Phi$-variables of the associate BasicBlock will be part of the dimensions
	of the polyhedron.

\begin{figure}[!h]
\centering
\begin{tikzpicture}
	\tikzstyle{bb} = [text width=6em, minimum height=12em]
	\tikzstyle{part}=[draw, fill=blue!10, text width=5em, 
    text centered, minimum height=2.5em, rounded corners, dashed]
	\tikzstyle{abstract}=[draw, circle, fill=red!20, text width=5em, 
    text centered, minimum height=2em,
	decorate,
    decoration={random steps,
		segment length=6pt,
		amplitude=4pt}
	]
    \node (bb) [bb]  {};
    \node (Phi) [part] [above of=bb] {\small $\Phi$-variables};
    \node (middle)  [below of=Phi] {};
    \node (inst) [part] [below of=bb] {\small Instructions};
	\node [abstract] (A) [left of=bb, xshift=-2cm] {Abstract \\ value};

	\path [draw, ->] (A) -- (middle);
	\path [draw, ->] (Phi) -- (inst);

    \path (inst.south)+(0,-1) node (label) {BasicBlock};

    \begin{pgfonlayer}{background}
        % Compute a few helper coordinates
        \path (Phi.north east)+(0.2,0.2) node (a) {};
        \path (inst.south west)+(-0.2,-0.2) node (b) {};
        \path[fill=yellow!20,rounded corners, draw=black!50, dashed]
            (a) rectangle (b);

        %\path (Phi.north west)+(-0.2,0.2) node (a) {};
        %\path (Phi.south east)+(+0.2,-0.2) node (b) {};
        %\path[fill=blue!10,rounded corners, draw=black!50, dashed]
        %    (a) rectangle (b);
        %\path (inst.north west)+(-0.2,0.2) node (a) {};
        %\path (inst.south east)+(+0.2,-0.2) node (b) {};
        %\path[fill=blue!10,rounded corners, draw=black!50, dashed]
        %    (a) rectangle (b);
       % \path (IR.north west)+(-0.2,0.2) node (a) {};
       % \path (IR.south -| IR.east)+(+0.2,-0.2) node (b) {};
       % \path[fill=blue!10,rounded corners, draw=black!50, dashed]
       %     (a) rectangle (b);
    \end{pgfonlayer}
\end{tikzpicture}
\end{figure}
\FloatBarrier

@@

	\begin{definition}[Liveness by linearity]
	A variable $v$ is \emph{live by linearity} at the control point $p$ iff:
	\begin{itemize}
	\item 
	its value is available at $p$, meaning that there is a path from the
	point defining the variable to $p$,
	\item one of these conditions holds:
		\begin{itemize}
		\item $v$ is live in $p$.
		\item There is a variable $v'$, defined as a linear combination of other
		variables $v_1, v_2, \dots, v_n$, such that $\exists i \in \{1,\dots,n\}, v = v_i$,
		and $v'$ is live by linearity in $p$.
		\end{itemize}
	\end{itemize}
	\end{definition}

	Finally, a variable is a dimension of $X_p$ iff it is live by linearity and
	it is not defined as a linear combination of program variables.

	There is a special case for $\Phi$-variables: in some cases, a
	$\Phi$-variable can be considered as a linear combination of program
	variables, when only a single incoming value is possible. This is the case
	when all but one polyhedra associated to a predessor are at bottom.
	For instance, let say point $p$ has $p_1, p_2, \dots, p_n$ as ordered 
	predecessors. At point $p$, we have $x = \Phi(x_1, x_2,\dots,x_n)$.
	If $\forall i \in \{2,\dots,n\}, X_i = \perp$, and $X_1 \neq \perp$, then we
	can use the linear relation $x = x_1$ instead of the $\Phi$ function. In
	practical cases, this is a way to reduce the size of the abstract values.
